Creating More Meaningful Explanations for Audio Deepfake Detection

### 2025-02-12 draft.tex
improve introduction

	In this work, we will make a distinction of two main categories of input features for a
	synthetic voice detection model: perceptible and imperceptible features. Perceptible features
	are features that can be perceived by the human ear, often vocal qualities such as jitter,
	shimmer or pitch fluctuation that have a wide range of uses even outside of audio
	classification such as diagnosis of disease \cite{chaiwongyen_deepfake-speech_2023}, while
	imperceptible features are typically out of the range of human hearing, and may not be
	directly reflect a vocal quality, otherwise referred to as "speaker-independent" features
	\cite{liu_hidden--wave_2023}. We use a new combination of features for our method, with an
	emphasis being placed on perceptible features, as we believe these to be the most likely to be
	understood by the layperson.

@@

	We then combine these features together using a novel model architecture, with separate
	sub-models for each input feature that are combined and processed in a final model, referred
	to as the terminus model, allowing us to process different kinds of features separately and
	avoid unnecessary, costly preprocessing.

@@

	Explanations are generated using the LIME method \cite{ribeiro_why_2016}, which allows us to,
	for an individual classification assess the impact of each input feature on the final result.
	In order to assess how useful the average local explanation is to a potential end user, and
	due to a lack of such assessments in previous research, we also introduce two metrics for
	aggregate evaluation of large number of generated LIME explanations, which we use to make a
	conclusion about the value of the selected feature set for the purpose of sensible

@@

@@

@@

@@

@@

>>>>

	In this work, we make a distinction of two main categories of input features for a synthetic
	voice detection model: perceptible and imperceptible features. Perceptible features are
	features that can be perceived by the human ear, often vocal qualities such as jitter, shimmer
	or pitch fluctuation that have a wide range of uses even outside of audio classification such
	as diagnosis of disease \cite{chaiwongyen_deepfake-speech_2023}, while imperceptible features
	are typically out of the range of human hearing, and may not be directly reflect a vocal
	quality, otherwise referred to as "speaker-independent" features \cite{liu_hidden--wave_2023}.
	We use a new combination of features for our method, with an emphasis being placed on
	perceptible features, as we believe these to be the most likely to be understood by the
	layperson.

@@

	We introduce a hybrid deep-learning model, with the purpose of combining these
	multidimensional features into a single output. The hybrid model consists firstly of a
	component sub-model for each of the input features, in order to address the problem of varying
	dimensionality of the extracted features without the drawbacks of costly transformations in
	the pre-processing phase. The final part of our hybrid model is what we will refer to as the
	terminus model, a model that uses the concatenated outputs of the individual sub-models as
	inputs and distills down to a singular output with a classic multi-layer perceptron
	architecture, allowing for an effective combination of the features into a final singular
	classification result.

@@

	In order to introduce explainability into our method, we generate explanations using a method
	that first generated the intermediate output data from the sub-models and then applies the
	Local Interpretable Model-agnostic Explanations (LIME) \cite{ribeiro_why_2016} on the terminus
	model, which allows us to, for an individual classification, assess the impact of each input
	feature on the final result through the local explanation, as, due the independence of the
	sub-models from one another, the importance of their outputs directly correlates with the
	importance of their inputs when it comes to the end-classification made by the terminus model.
	We then stood before the problem of assessing our explanations, more specifically, assessing
	the impact of the input features on the average local explanation, in order to make a
	conclusion about the usefulness or promise of our chosen features for providing understandable
	explanations of our model's classifications. In order to tackle the problem with our local
	explanations, and due to a lack of metrics to assess the aggregated average value of local
	explanations, we also introduce two metrics for such an aggregate evaluation of large number
	of generated LIME explanations, importance and trust, on a per-feature basis, which we use to
	make a conclusion about the value of the selected feature set for the purpose of sensible

@@

	The contributions of this work are in summary: Our hybrid model-architecture and modified
	LIME-based method for generating explanations of classifications of audio voice samples as
	synthetic or bona-fide and the two metrics that we introduce in order to assess the impact of
	individual features over numerous local explanations.
	This paper is divided into the following sections:
		\item §\ref{sec:related} Related Work, where we discuss the most relevant related
			research in the field, in order to create a picture of the current state-of-the-art
		\item §\ref{sec:method} Method, where we describe the general model implementation, as
			well as the implementation of our LIME-based method for generating explanations in
			more detail.
		\item §\ref{sec:experimental_results} Experimental Results, where we discuss how we performed our
			experiments, how our hybrid model performs, as well as use our proposed metrics in
			order to assess the feature impact in the average explanation.
		\item §\ref{sec:discussion} Discussion, where we discuss the implication of the results of
			our experiments, including the potential of the explanations as well as the
			shortcomings and limitations of our method.

@@

	\label{sec:related}

@@

	\label{sec:method}

@@

	\label{sec:experimental_results}

@@

	\label{sec:discussion}
